Fast Random Integer Generation in an Interval

### 2024-02-25 Presentation/main.tex
remove FDR + summary table (need to fix vertical line)
table fix
finished presentation v1
new motivation + better [a,b) visual + bit representation + further fixes

\newcommand{\RNW}{\ensuremath{\mathcal{R}_n^W}}

@@

@@

\tikzset{
	legend/.style = {font=\boldmath, align=left},
	every path/.style = {line width=.57mm, line cap=round},
}

@@

>>>>

\newcommand{\RNW}{\ensuremath{\mathcal{R}}}

\usepackage{array}
\newcolumntype{P}[1]{>{\centering\arraybackslash}p{#1}} % Centered Multilines
\usepackage{booktabs}
\usepackage{pifont} % checkmark / cross - thanks to `https://tex.stackexchange.com/questions/42619/xmark-that-complements-the-ams-checkmark`
\newcommand{\cmark}{\ding{51}}
\newcommand{\xmark}{\ding{55}}

@@

\usepackage[mode=buildnew]{standalone}

@@

\tikzset{
	legend/.style = {font=\boldmath, align=left},
	every path/.style = {line width=.57mm, line cap=round},
	time/.style={very thick,|-latex, gray},
	int/.style={ultra thick,{|[scale=.5]}-{|[scale=.5]}},
}

@@

\usepackage{xparse}

% Thanks to `https://tex.stackexchange.com/questions/8720/overbrace-underbrace-but-with-an-arrow-instead`
\renewcommand{\arraystretch}{0.7}
\NewDocumentCommand{\overarrow}{O{0} O{\big\downarrow} m}{%
  \overset{\makebox[0pt]{\begin{tabular}{@{}c@{}}$#3$\\[0pt]\ensuremath{#2}\\[0pt]\hfil\end{tabular}}}{#1}
}

### 2024-02-25 Presentation/preliminaries.tex
remove FDR + summary table (need to fix vertical line)
some fixes
finished presentation v1
new motivation + better [a,b) visual + bit representation + further fixes

\subsection{Formal Definition}
\begin{frame}{Formal Definition}
    \pause
    \onslide<+->{Setting:}
    \begin{itemize}[<+->]
        \item \textbf{Input:} upper bound of interval $n \in \mathbb{N}$
        \item \textbf{Output:} uniform random integer in interval $[0,n)$
    \end{itemize}

@@

    \begin{alertblock}{}<+->
        \centering
        But what if we want a random integer in $[a,b)$ for $a,b \in \mathbb{N},\,0 < a < b$ instead?
    \end{alertblock}

    \smallskip

    \onslide<+->{We can map this to our setting by subtracting $a$!}
    \begin{itemize}[<+->]
        \item Set $n = b - a$ and draw a uniform random integer $x \in [0,n)$
        \item Return $x + a$
    \end{itemize}
\end{frame}

@@

\subsection{The Naive Approach}
\begin{frame}{The Naive Approach}
    \pause 
    \begin{block}{How do we get random numbers?}<+->
        \begin{itemize}[<+->]
            \item Generated by Pseudo-Random-Number-Generators (PRNGs)
            \item Generated as $W$-bit words, i.e. unsigned integers in $[0,2^W)$ (typically $W \in \{32,64\}$)
        \end{itemize}
    \end{block}

    \onslide<+->{\begin{center}
            $\texttt{rand()} \Mod n$
    \end{center}}

@@

\begin{frame}{The Naive Approach}
    \pause 
    \onslide<+->{In general, applying $x \Mod n$ to $[0,2^W)$ yields}
    \onslide<+->{\begin{align*}\everymath{} % Allows specific coloring
        \overbrace{\underbrace{\overbrace{0,1,\ldots,n - 1}^{\text{$n$ values}},\overbrace{0,1,\ldots,n - 1}^{\text{$n$ values}},\ldots,\overbrace{0,1,\ldots,n - 1}^{\text{$n$ values}}}_{\text{$\left(2^W \div n\right) \cdot n$ values}},\ER{\underbrace{0,1,\ldots,\RNW - 1}_{\text{$\RNW$ values}}}}^{\text{$2^W$ values}}
    \end{align*}}

@@

    \onslide<+->{\GB{Every} approach that maps every integer in $[0,2^W)$ to a single number in $[0,n)$}
    \onslide<+->{does \ER{not} generate uniform random integers in one step}
    \onslide<+->{whenever $n$ does not divide $2^W$.}

    \onslide<+->{\GB{Idea:} Use \ER{rejection sampling} to achieve uniformity!}
\end{frame}

>>>>

\subsection{Formal Definition}
\begin{frame}{Formal Definition}
    \pause
    \onslide<+->{\textbf{Input:}} \begin{itemize}[<+->]
        \item source of uniform random integers in $[0,2^W)$: \texttt{rand()}
        \item upper bound of interval $n \in \mathbb{N}$
    \end{itemize}
    \onslide<+->{\textbf{Output:} \begin{itemize}
        \item uniform random integer in interval $[0,n)$
    \end{itemize}}

@@

    \begin{center}
        \begin{tikzpicture}[scale = 1.5]
            \draw<6->[time] (0,0) -- (9,0);
            \node<6->[time] at (0.015,-0.25) {$0$};

            \draw<7->[int, emorot] (5,0) -- (8,0);
            \node<7-> at (5.015,-0.25) {$\ER{a}$};
            \node<7-> at (8.0,-0.25) {$\ER{b}$};
            \node<7-> at (6.5,0.35) {$\ER{[a,b)}$};

            \draw<8->[int, goetheblau] (0,0) -- (3,0);
            \node<8-> at (3.0,-0.25) {$b - a$};
            \node<8-> at (1.5,0.35) {$[0,b - a)$};

            \draw<9->[time,black] (3.3,0.5) -- (4.7,0.5);
            \node<9-> at (4.0,0.72) {$\BB{+\,a}$};

            \node[opacity=0] at (4.5,1.0) {};
        \end{tikzpicture}
    \end{center}
\end{frame}

@@

\subsection{The Naive Approach}\label{sec:1.3}
\begin{frame}{The Naive Approach}
    \pause 
    \onslide<+->{\begin{center}
            \large{$\texttt{rand()} \Mod n$}
    \end{center}}

@@

\begin{frame}{The Naive Approach - Bias}
    \pause 
    \onslide<+->{In general, applying $x \Mod n$ to $[0,2^W)$ yields}
    \onslide<+->{\begin{align*}\everymath{} % Allows specific coloring
        \overbrace{\underbrace{\overbrace{0,1,\ldots,n - 1}^{\text{$n$ values}},\overbrace{0,1,\ldots,n - 1}^{\text{$n$ values}},\ldots,\overbrace{0,1,\ldots,n - 1}^{\text{$n$ values}}}_{\text{$\left(2^W \div n\right) \cdot n$ values}},\ER{\underbrace{0,1,\ldots,\RNW - 1}_{\text{$\RNW$ values}}}}^{\text{$2^W$ values}}
    \end{align*}}

@@

    \begin{block}{Deterministic Mappings}<+->
        \onslide<+->{\GB{Every} deterministic mapping $f\colon[0,2^W) \rightarrow [0,n)$}
        \onslide<+->{does \ER{not} generate \GB{uniform} random integers in one step}
        \onslide<+->{whenever $n$ does not divide $2^W$.}
    \end{block}
\end{frame}
